\section{Needs Analysis}
A thorough needs analysis is critical in software development. Any ambiguity, contradiction, or missing information found in the
requirements document can lead to significant setbacks, potentially requiring a restart of the project. In this section, we will
take a detailed look at how the requirements document is structured.

\begin{prettyBox}{Note}{red} 
\textbf{Difference Between Goals \& Needs ?}\\

It is crucial to understand that the requirements document holds the client's needs and not their goals . A goal is subjective and
open to interpretation, while a need is objective—measurable or verifiable.\\

\textbf{\underline{Example :}}\\

The client might request a "pleasant user interface" , This can be interpreted in many ways:
\begin{itemize}
    \item a visually appealing UI with lots of colors and animations   
    \item an easy-to-use interface
    \item a minimalistic design
\end{itemize} 

and so on. To turn this into a clear need, we must clarify what the client means. For instance, they might want the UI to be 
organized with all features accessible through a dropdown menu.

\end{prettyBox}

\subsection{Requirments Document Structure}
\begin{itemize}
   \item \textbf{Introduction : }This section provides an overview of the document by outlining its key components :
      \begin{itemize} 
          \item Purpose: The reason for creating this document, which is primarily to align the development team and facilitate
           communication with the client.
           \vspace{0.15cm}
          \item Scope: A high-level summary of the software’s main functionality, without going into too much detail. 
           \vspace{0.15cm}
          \item Context: The reason for creating the software, which could be to sell it to a client or company, to develop an
          open-source project, or other similar motivations.\\
      \end{itemize}    
    \item \textbf{Hardware : }This section states whether the software requires any special hardware components like : GPU,
sensors, or a camera ...etc . It also outlines the minimum hardware requirements needed to run the software, as well as the optimal hardware configuration for the best and smoothest experience.
    \item \textbf{Conceptual Model : }This section describes the overall software architecture through a high-level graphical
        representation, highlighting key components and their relationships.\hspace{0.1cm}It provides an overview of how the system is structured and
operates, making it easier for stakeholders to understand.

\textbf{\underline{Example :}}

A desktop system composed of an email service, a spreadsheet, a document processing service, and an information retrieval service.
\end{itemize}
\begin{center}
\begin{tikzpicture}
    \draw (0,0) rectangle (2,1);
    \node at (1,0.5){User};
    \draw[->] (1,1) -- (1,2) -- (10.5 ,2) -- (10.5,-2);
    \draw[->] (2,0.5) -- (6.25,0.5) -- (6.25,0);
    \draw[->] (0,0.5) -- (-1.5,0.5) -- (-1.5,-8) -- (1.75,-8) -- (1.75,-7);
    \draw[->] (1.625,0) -- (1.625,-1);

    \draw (0.5,-3) rectangle (2.75,-1);
    \node at (1.625,-1.75) {Email};
    \node at (1.625 ,-2.25){Management};
    \draw[->] (1.625,-3) -- (1.625,-4);
    \draw[->] (0.5,-2) -- (-0.5,-2) -- (-0.5,-6.5) -- (0.75,-6.5);
    \draw[->] (2.75,-2) -- (4.125,-2) -- (4.125,-1) -- (5.25,-1);

    \draw (0.75,-5) rectangle (2.75,-4);
    \node at (1.75,-4.5) {Network};

    \draw (0.75,-7) rectangle (2.75,-6);
    \node at (1.75,-6.5) {Spreadsheet};
    \draw[->] (2.75,-6.5) -- (5.25,-6.5);
      
    \draw (5.25,-7) rectangle (7.25,-6);
    \node at (6.25,-6.5) {DB};

    \draw(5.25,-2) rectangle (7.25,0);
    \node at (6.25,-0.75) {Document};
    \node at (6.25,-1.25) {Processing};
    \draw[->] (6.25,-2) -- (6.25,-6); 
  
    \draw (9.5,-4) rectangle (11.5,-2);
    \node at (10.5,-2.75) {Information};
    \node at (10.5,-3.25) {Retrieval};
    \draw[->] (10.5 , -4) -- (10.5,-6.5) -- (7.25,-6.5); 

 \end{tikzpicture}
\end{center}

\vspace{0.75cm}
The next step consist into creating a new conceptual model for each complexe fonction
\vspace{0.25cm}
\begin{center}
\begin{tikzpicture}
    \draw (-3,0) rectangle (-1,1);
    \node at (-2,0.5){User};
    \draw[->] (-2,0) -- (-2,-2.625) -- (0.75,-2.625);
    \draw[->] (-1.75,0) -- (-1.75,-0.625) -- (0.75,-0.625);
    \draw[->] (-2.25,0) -- (-2.25,-4.625) -- (0.75,-4.625);
    
    \draw (0.75,-1.25) rectangle (2.75,0);
    \node at (1.75,-0.475){Message};
    \node at (1.75,-0.875){Creation};
    \draw[->] (2.75,-0.625) -- (4,-0.625) -- (4,-6.75) -- (2.75,-6.75);
    \draw[->] (2.25,0) -- (2.25,0.5) -- (6.25,0.5) -- (6.25,0);
    \draw[->] (1.75,0) -- (1.75,1) -- (9.5,1) -- (9.5,-0.5);

    \draw (0.75,-3.25) rectangle (2.75,-2);
    \node at (1.75,-2.475){Message};
    \node at (1.75,-2.875){Deletion};
    \draw[->] (2.75,-2.625) -- (3.25 ,-2.625) -- (3.25,-6.5) -- (2.75,-6.5);

    \draw (0.75,-5.25) rectangle (2.75,-4);
    \node at (1.75,-4.475){Message};
    \node at (1.75,-4.875){Transfer};
    \draw[->] (1.75,-5.25) -- (1.75,-6);

    \draw (0.75,-7) rectangle (2.75,-6);
    \node at (1.75,-6.5) {Network};

    \draw (8.5,-1.5) rectangle (10.5,-0.5);
    \node at (9.5,-1) {Spreadsheet};
    
    \draw(5.25,-2) rectangle (7.25,0);
    \node at (6.25,-0.75) {Document};
    \node at (6.25,-1.25) {Processing};

\end{tikzpicture}

\end{center}
\begin{itemize}
\item \textbf{Functional Requirements:} These define what the system should do, focusing on the specific features and functions the
software must deliver to meet user or business needs. They describe the expected behavior of the system in various situations.
Functional requirements are concrete and measurable. They can be expressed using natural, semi-formal, or formal language,
or a mix of these. 

\begin{itemize} 
\item \textbf{Natural Language:} Easy to implement and understand, but lacks structure and precision, which can lead to ambiguity.
It makes automating analysis of the document harder, relying heavily on the writer’s linguistic experience. 
\item \textbf{Structured (Semi-Formal) Language:} Limited use of natural language, more structured and precise than natural language
, often accompanied by graphical notations. 
\item \textbf{Formal Language:} Hard to master and time-consuming to implement. It is difficult for clients to understand but is based
on mathematical theory, making it the most precise language and easier to automate verification. 
\end{itemize}

\item \textbf{Non-Functional Requirements:} Define the restrictions and constraints related to both hardware and software within
the context of the ongoing project. Non-functional requirements are particularly influenced by changes in technologies 
(both hardware and software) and are crucial for complex software systems. As the project develops, changes in hardware may occur.
These changes can be anticipated by projecting the expected performance levels that will be required by the end of the project.
\item \textbf{Maintenance Information:} Anticipates possible actions after the software's initial release, such as adding
new features, improving performance, or addressing potential issues.
   \item \textbf{Glossary:} Provides definitions of the terms and concepts used in the document to help readers. This ensures
that the terminology is clear, as the requirements document is shared and read by the design team, developers, and stakeholders, 
without assuming prior knowledge of these terms.
  \item \textbf{Index:} Helps the reader find specific topics and sections of the document more efficiently by providing a
detailed list of references to relevant sections, parts, and page numbers.
\end{itemize}


\begin{prettyBox}{Note}{red}
 Functional and non-functional requirements are inherently connected, and their interplay can sometimes lead to conflicts. 
 By understanding the potential for these conflicts and establishing a process for managing them, development teams can better
 balance user needs and system performance, ultimately leading to a more successful product.
\end{prettyBox}

\subsection{Requirements Validation}
The requirements need to be coherent, realizable, and complete. Anticipation of hardware needs to be considered. 
It is crucial for the requirements document to be validated in order to initiate the next steps of the software life cycle.

\begin{itemize}
    \item review Technique is an efficient way to monitor and update the requirements .
    \item There are various analysis tools available that can facilitate the validation process of the requirements document,
helping to ensure accuracy and completeness.
\end{itemize}

