
\section*{\textit{\underline{Exercise 2:}} "Choosing an IT Solution"}
A research laboratory wants to equip itself with an integrated management system for its research teams to handle information
related to researchers, their scientific productions in conference papers, the number of journal articles published, the number 
of research trips, and scientific leave.

To achieve this, three possible approaches are available to the laboratory: 
\begin{enumerate}
    \item Acquire a database management system and develop its own system based on this tool. 
    \item Purchase a similar system from another laboratory and adapt it to its own needs. 
    \item Join a group of laboratories, establish a common requirements document, and contact a software vendor to develop a
single system for all. 
\end{enumerate}

\textbf{Question:} Identify two possible risks for each of these strategies and propose risk resolution techniques that would help
decide which approach to adopt.

\vspace{1.5cm}

{\textbf{\underline{\large{Solution} :}}\\

Risks of each solution :

\vspace{0.5cm}
\hspace{0.5cm}1st Solution 
\begin{itemize}
    \item Risk of insufficient budget to acquire a database management system.
    \item Potential development challenges due to lack of expertise.
    \item Risk of extended development time.
\end{itemize}
}

\hspace{0.5cm}2nd Solution 
\begin{itemize}
    \item Risk of insufficient budget to purchase the software.
    \item Challenges in adapting the software due to limited expertise or lack of customizability. =\textgreater maybe 
acquiring new software that meets the lab's requirements is better than to adapt an existing one.
\end{itemize}

